---
title: "Assignment 1"
author: "Neo Moloi: 202000784"
date:  "2020/11/18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 BASICS
## 1.1. Variables in R
What is the value of " B " after running these command? [2]
```{r}
a <- 5.0
c <- 10
t <- 12
B <- a+c/t

```
The value of B is: ``r round(B,2)``

What type of vector do you get in each of the following? [4]
```{r}
vec_1 <- c(44, 5, 0, "k")
vec_2 <- c(1, 2, 3, FALSE)
vec_3 <- c("a", "b", "c", TRUE)
vec_4 <- c(1, 2, 3, "4")
```
The type of vector for vec_1 : ``r typeof(vec_1)``

The type of vector for vec_2 : ``r typeof(vec_2)``

The type of vector for vec_3 : ``r typeof(vec_3)``

The type of vector for vec_4 : ``r typeof(vec_4)``

## 1.2. Conditional subsetting
Consider the following vector:
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)

```
### 1.2.1. Create a vector vec_1 with no missing values using conditional subsetting and explain your code [3]
The code:
```{r}
Vec_1<-na.omit(vec)
print(Vec_1)
```
The output:

``r na.omit(vec)``

Explanation of code:

Using the `` na.omit()`` R function we can remove all incomplete objects in a data object. In this case the vector is the data object. The 'NA' values are the incomplete/missing values. The code will remove every instance of incomplete/missing values.Every 'NA' will be removed. 

### 1.2.2. Create a vector of indexes of missing values using conditional subsetting and explain your code [3]
The code :
```{r}
vec[c(FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,TRUE)]
```


The codes output: 

``r vec[c(FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,TRUE)]``

Explanation of code:

 Using a logical vector. `TRUE` will select the element with the same index, while `FALSE` will not.In this case the `FALSE` has been allocated to complete objects and `TRUE` has been allocated to incomplete objects.It will select all the objects with "NA".
 
### 1.2.3. Write a code to determine how many missing values are in vec. [2]
The code: 

```{r}
length(vec[c(FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,TRUE)])
```

The output:

``r length(vec[c(FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,TRUE)])``

### 1.2.4. Write a R code to create a data frame which contains details of 3 students as follows: [4]
The code:
```{r}
Students = data.frame(Name=c("Neo","Gary","Onthatile"),
                      Surname=c("Moloi","Leander","Seodigeng"),
                      Age=c(19,22,18),
                      Gender=c("Male","Male","Female")
                      )
print(Students)
```

# 2. Functions in R
## 2.1. Create a function which returns a mean of an input vector and explain your code [4]
The code:
```{r}
 example_1 <-c(2,4,NA,6,10,NA,12,14)
 example_2 <-c(1,NA,3,5,7)

mean_value <- function(input) {
  output<-mean(input, na.rm = TRUE)
  return(output)
}
```
The output:

The mean value for example_1 is :``r mean_value(example_1)``

The mean value for example_2 is :``r mean_value(example_2)``

Explanation of code:

Define a function `mean_value` which gives us the mean value of a given vector.After defining the function we  assign it to the output of function. The function will have take one argument which are the vectors, `example_1 ` and `example_2`. A variable called `output` is assigned to a `mean` function.The `mean` function looks at the values of the vectors and `na.rm = TRUE` removes the missing values from the input vector.The `mean` adds all the objects in the vector and divides it by its length to get the mean value.



## 2.2. Create a function which take no inputs , it reads a Titanic .csv file and returns its content and explain your code. [3]
The code:
```{r}
library(tidyverse)
df3 <-tibble(read.csv("train.csv"))
print(df3)


```

Explanation of code:


# 3. Conditionals and Loops
## 3.1. Write a function that take an input below and returns a transformed vector below into 1 when its element is male and 2 when its female using ifelse. [8]
```{r}
g <- c("MALE","FEMALE","FEMALE","MALE","FEMALE","MALE","MALE","FEMALE", "MALE","MALE","FEMALE","MALE","MALE",
"FEMALE","MALE","MALE","MALE")
```
The code:
```{r}
Gender <-function(input){
  output<-ifelse(input == "MALE",1,2)
  return(output)
}
```
The output:

`r Gender(g)`

Explanation of code:

Define a function called `Gender` which returns a transformed vector into `1` when its element is `MALE` and `2` when its `FEMALE` using `ifelse`.After defining the function we  assign it to the output of function. The function will have take one argument which is the vector`g`.A variable called `output` is assigned to a `ifelse` function. The test expression is the objects in the `g`. If the object in the vector `g` is `Male` it will return `1` else it will return `2`.

## 3.2. Generate x from normal distribution and sample with replacement "M" and "F" for male and female respectively by running the code below.
```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
```
### 3.2.1. Use the "cbind " function in R to create one dataset and assign it to df. [2]
The code:
```{r}
df <- cbind(x,y)
print(head(df))
```

### 3.2.2. Convert df from 3.2.1 into a data.frame [2]
The code:
```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)

df2 <-data.frame(df)
print(head(df2))
```

### 3.2.3. Write a code using tidyverse to compute a number of males ("M") and females ("F") in 3.2.1 [3]
The code:
```{r}
library(tidyverse)

df2 %>% group_by(y) %>% count
```


### 3.2.4. Create a bar plot using ggplot to visualize your results in 3.2.3 [5]
The code:
```{r}
library(ggplot2)

df2 %>% group_by(y) %>% count %>% 
  ggplot(aes(x=y,y=n))+
  geom_col()+
  labs(y= "Number of people", x= "Gender",title = "Relationship between gender and a random number")
   
```


### 3.2.5. Use mutate function to add sum of x per "M" and "F" and plot the results using sum of x per "M" and "F". [5]
The code:
```{r}
library(ggplot2)
df2 %>% group_by(y) %>% count %>% 
 ggplot(aes(x=y,y=n))+
  geom_col()+
  labs(y= "Number of people", x= "Gender",title = "Relationship between gender and a random number")
```

# THE END OF THE ASSIGNMENT

          
